---
title: "Final Project"
author: "Al Haque"
date: "5/14/2022"
output: 
  html_document:
    theme: cerulean 
    toc: true
    toc_float: true
    
---

 My final project is my attempt at creating a basic recommender system  using the MovieLens dataset I've found on Kaggle. This is my process at exploring the data,preprocessing the data,creating the rating matrix,building a model and implmenting the model to get a recommendations of movies per user. 





```{r include=TRUE}
library(tidyverse)
library(recommenderlab)
library(stringi)
library(reshape2)

```

```{r}
## insert the MovieLens Dataset: 

ratings <-read.csv("https://raw.githubusercontent.com/AldataSci/FinalProject/main/ratings.csv",header=TRUE)
movies <- read.csv("https://raw.githubusercontent.com/AldataSci/FinalProject/main/movies.csv",header=TRUE)
```

------

## Data Exploration: 

```{r}
## There are various rating from 0.5 to 5 stars rating.. 
vector_rating <- as.vector(unique(ratings$rating))
table_rating <- table(vector_rating)
table_rating

```


```{r}
## Looking at the distribution of the rating.. 
factor_vector <- as.factor(ratings$rating)
qplot(factor_vector) + ggtitle("Distribution of the Rating") +
  theme_bw()
```
```{r}
## Exploring which movies have the most views or most rating! in this small dataset!

## arrange the movie ratings that have been rated the most
movie_Rating <- ratings %>%
  group_by(movieId) %>%
  count() %>%
  arrange(desc(n))

## do an inner join to get see which movies has been reviewed
Movie_rating <- movie_Rating %>%
  inner_join(movies,by="movieId") %>%
  select(-genres)

ggplot(Movie_rating[1:10,],aes(x=title,y=n))+
  geom_bar(stat="identity") + 
  ggtitle("Top Ten Movies That Were Given A Rating") +
  xlab("Movie") +
  ylab("Movie Rated By A User") +
  coord_flip()
```
```{r}
## Exploring Average Rating Per Movie: 

avg_movie_rating <- ratings %>%
  select("userId","movieId","rating") %>%
  group_by(movieId) %>%
  summarise(avg_rating = mean(rating)) %>%
  inner_join(movies,by="movieId") %>%
  select(-genres)


## left-tail skewed again and some prominent peaks but we can see that most of the average ratings per movies are around a 3 or a 4 maybe we should get rids of 0.5 and 1 to get a better distribution of the data.. 

ggplot(avg_movie_rating,aes(x=avg_rating)) +
  geom_histogram(color="darkblue",fill="lightblue") +
  theme_classic() +
  ggtitle("Distribution of the Average Movie Rating")
```

```{r}
## Here I filtered out the average rating greater than 3 and less than 4.5 to have less skewed datas.. 
avg_movie_rating_clean <- ratings %>%
  select("userId","movieId","rating") %>%
  group_by(movieId) %>%
  summarise(avg_rating = mean(rating)) %>%
  filter(avg_rating > "2.5" & avg_rating <= "4.5") %>%
  inner_join(movies,by="movieId") %>%
  select(-genres)


## Distribution Looks Okay when we removed the exteremes but there are too many dramatic peaks in the datas..
## we removed this since possibly a few people have accounted for the average rating.. 
ggplot(avg_movie_rating_clean,aes(x=avg_rating)) +
  geom_histogram(color="purple",fill="lightblue") +
  theme_classic() +
  ggtitle("Distribution of the Average Movie Rating With Low Rating Removed")
```

-------

## Data Preprocessing: 


```{r}
## In order to create a rating matrix I had to convert the datings where each row was a user and each column was a movieId and a cell was a rating of a user given to that particular movie

ratingmat = dcast(ratings, userId~movieId, value.var = "rating", na.rm=FALSE)

dim(ratingmat)

```



```{r}
## remove the UserId:
ratingMat <- ratingmat %>%
  select(-userId) %>%
  as.matrix()
RatingMat <- ratingMat %>%
  as("realRatingMatrix")

## visualize the Heatmap and we can see that the matrix is mainly sparse.
## A rating matrix of the first 50 users and 50 movie ratings. 
image(ratingMat[1:50,1:50], main ="Heatmap of the Rating Matrix")


```


## Building A Model:

```{r}
## spliting the data into a testing and training model 
training <- sample(x=c(TRUE,FALSE),size=nrow(RatingMat),replace=TRUE,prob=c(0.8,0.2))
recc_training_data <- RatingMat[training,]
recc_testing_data <- RatingMat[!training,]

```



```{r}
## Build a UBCM since we are making the recommender based upon the users averages to recommend movies
## use the default method of cosine.


recc_model <- Recommender(data=recc_training_data,method="UBCF",parameter=list(method="Cosine"))

## use the predict model to make prediciton

Top_5_pred = predict(recc_model, recc_testing_data, n=5)

## define a dataframe having a column with the ordered movie labels, access the movie names with the item Label slots..
movie_Labels <- data.frame(movieId = Top_5_pred@itemLabels)
movie_Labels <- movies %>%
  left_join(movies,by="movieId")

```
## Implementing The Model: 

```{r}
## look at the recommendation of the first user: 
recc_user_1 <- Top_5_pred@items[[1]]
recc_user_1
movie_Labels$title.x[recc_user_1]
Top_5_List = as(Top_5_pred, "list")
```


```{r}

## Created A matrix to help show the numbers of users: 
recc_Matrix <- sapply(Top_5_pred@items,function(x){
  recommendation <- movie_Labels$title.x[x]
})

## Showing the Recommendation For the First 3 Users: 

recc_Matrix[,1:5]


```


```{r}

## We check if some movies are much more likely to be recommended. 
## visualize the movie distribution:
## Some movies have been recommended 10 times or fewer and a few has been recommended 20 or 24 times. 
recc_Matrix_table <- table(recc_Matrix)
recc_Matrix_Num <- as.numeric(recc_Matrix_table)

bin_recc <- cut(recc_Matrix_Num,breaks=c(0,10,20,max(recc_Matrix_Num)))

qplot(bin_recc) + ggtitle("Distribution of recomendations per Users")

```

------ 

### Sources/Acknowledgements: 

___1.Usuelli, Michele, and Suresh K. Gorakala. Building a Recommendation System with R. Packt Publishing Limited 2015.____

___2.Pierson, Lillian. “How to Build a Recommendation Engine in R: Marketing Data Science!” Data, 22 Mar. 2021, https://www.data-mania.com/blog/how-to-build-a-recommendation-engine-in-r/.____


__3.Efficient Reshaping Using Data.tables, 22 Sept. 2021, https://cran.r-project.org/web/packages/data.table/vignettes/datatable-reshape.html.____

-------

## Thank you! 
